Lyman-$\alpha$ photons through rotating outflows

### 2018-03-31 paper/paper.tex
references

Recent advances in instrumentation have revolutionized the kind of
kinematic studies that can be performed on extra-galactic systems with small angular
sizes on the sky, revealing the presence of gas rotation on vastly
different physical scales.
For instance, spatially resolved spectra on compact dwarf galaxies
have measured clear signs of gas showing pure rotation kinematics
\citep{2015A&A...577A..21C,2017A&A...600A.125C}; recent mapping of
high redshift circumgalactic regions have also revealed kinematic
evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
These systems with star formation, some neutral gas and low dust
contents naturally show a \lya emission line \citep{PartridgePeebles},
motivating the observational work to phenomenologically link
H$\alpha$ kinematics to \lya spectra \citep[e.g.][]{Herenz2016}.

Theoretically speaking, what is then the expected imprint of rotation
on a resonant emission line such as the \lya line? To what extent is
possible to constrain rotational kinematics from the \lya emission line? 
Detailed radiative transfer (RT) \lya modeling of rotating systems
started until recently by \cite{Garavito14}.
In that work the authors studied the influence of pure solid body
rotation on the \lya line's morphology.
They found that rotation indeed introduces changes, the most
noticeable being the dependence of the spectra with the viewing
angle with respect to the rotation axis.

>>>>

Recent advances in instrumentation have revealed the presence of gas
rotation on vastly different physical scales.
For instance, spatially resolved spectra on compact dwarf galaxies
have measured clear signs of gas showing pure rotation kinematics
\citep{2015A&A...577A..21C,2017A&A...600A.125C} and the recent mapping of
high redshift circumgalactic regions have also revealed kinematic
evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
These systems with star formation, neutral gas and low dust
contents naturally show a \lya emission line \citep{PartridgePeebles}
motivating the observational work to phenomenologically link
H$\alpha$ kinematics to \lya spectra \citep[e.g.][]{Herenz2016}.

What is then the expected imprint of rotation on a resonant emission
line such as the \lya line? To what extent is possible to constrain
rotational kinematics from the \lya emission line? 
Detailed radiative transfer (RT) \lya modeling of rotating systems
started until recently by \cite{Garavito14}.
In that work the authors studied the influence of pure solid body
rotation on the \lya line's morphology.
They found that rotation indeed introduces changes, the most
noticeable being the dependence of the spectra with the viewing
angle with respect to the rotation axis.
